\newcommand{\pubyear}[1]{#1}
\documentclass[master,bottom,nosig]{usbthesis}
% \usepackage{geometry}
%  \geometry{
%  a4paper,
%  total={8.5in,11in},
%  left=1in,
%  right=1in,
%  top=1in,
%  bottom=1in,
%  }
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{braket}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{physics}
\usepackage{tikz-cd}
\usepackage{musicography}
\usepackage{enumerate}
\usepackage{xfrac}
\usepackage[parfill]{parskip}
\usepackage{subcaption}
\usepackage{float}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{color}
\usepackage{bm}
\usepackage{setspace}
\usepackage{array}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{afterpage}
\usepackage{hhline}
\usepackage{url}
%\usepackage{indentfirst}
\usepackage{footmisc} % provides \footref and \mpfootnotemark commands
\usepackage{xspace}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{pdfpages}
\usepackage[nottoc]{tocbibind}
\hyphenation{put words here which LaTeX does not hy-phen-ate pro-per-ly}
\usepackage{pgfplots}
\pgfplotsset{compat=1.12}
\usepgfplotslibrary{fillbetween}

% \newcommand{\cm}{\mathbb{C}}
% \newcommand{\re}{\mathbb{R}}
% \newcommand{\dvg}{\nabla\cdot}
% \newcommand{\crl}{\nabla\times}
% \newcommand{\proj}[1]{\ensuremath{{\ket{#1}\bra{#1}}}}
% \newcommand{\vect}{\text{Vect}}
% \newcommand{\smooth}{C^{\infty}}
% \newcommand{\dif}{\text{d}}
% \DeclareMathOperator{\aext}{\hspace{2pt}\lrcorner\hspace{2pt}}

% \makeatletter
% \newcommand*\bigcdot{\mathpalette\bigcdot@{.8}}
% \newcommand*\bigcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
% \makeatother
% \newlength{\arrow}
% \settowidth{\arrow}{\scriptsize$1000$}
% \newcommand*{\myrightarrow}[1]{\xrightarrow{\mathmakebox[\arrow]{#1}}}
% \newcommand*{\TakeFourierOrnament}[1]{{%
% \fon`'tencoding{U}\fontfamily{futs}\selectfont\char#1}}
% \newcommand*{\danger}{\TakeFourierOrnament{66}}
% \setlength{\jot}{1.5mm}

\theoremstyle{definition}
\newtheorem{thm}{Theorem}[section] % the main one
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{joke}{Joke}
\newtheorem{qn}{Question}
% other statement types

% for specifying a name
\theoremstyle{plain} % just in case the style had changed
\newcommand{\thistheoremname}{}
\newtheorem{genericthm}[thm]{\thistheoremname}
\newenvironment{namedthm}[1]
  {\renewcommand{\thistheoremname}{#1}%
   \begin{genericthm}}
  {\end{genericthm}}
  
\newtheorem{defn}{Definition}
\newtheorem{prop}{Proposition}
\newtheorem{rmk}{Remark}
\newtheorem{conj}{Conjecture}
%\newtheorem{lemma}{Lemma}
\newtheorem{cor}{Corollary}
\newtheorem{problem}{Problem}

\author{Evan Saraivanov}%
\title{Accelerating Cosmological Analysis Using Neural Networks}%

\month{May}
\pubyear{2023}%
\program{Physics}%
\director{Vivian Miranda}{Assistant Professor, Deparment of Physics and Astronomy, Yang Institue for Theoretical Physics}%
\chairman{Anja von der Linden}{Associate Professor, Department of Physics and Astronomy}%
\fstmember{Navid Vafaei-Najafabadi}{Assistant Professor, Department of Physics and Astronomy}%
\outmember{\,}{\,}{\,}%
\dean{\,}%

\begin{document}

\singlespacing %
\pagenumbering{roman} %
\maketitle %
\makeapproval %

\begin{abstract}
Despite many efforts, no cosmological model has been proposed that can lower the $H_0$ and $S_8$ tension and also agree with ongoing experiments. Fortunately, however, we are approaching a great age for cosmology. In the near future, many telescopes will begin to operate, taking more detailed data than ever before. Current methods of analyzing models on real data are slow, taking multiple days or even weeks to generate the necessary data. With the number of proposed models, analyzing each of them with traditional means is futile. This thesis presents one model independent extension to the accepted model of cosmology, $\Lambda$CDM, which splits the matter density into two components. This highlights the traditional analysis pipeline as well as presents a consistency check within cosmological analysis. This thesis then presents novel methods for using neural networks to learn the time-consuming parts of this pipeline, greatly accelerating the analysis of cosmological models. Finally, this new neural network design is applied to computing tension metrics between the upcoming LSST experiment and Planck 2018 data. Tension metrics are frequentist estimators, and thus have error bars. Using traditional means the computation of these errors is intractable, but using the novel neural network the errors can be estimated in a matter of weeks.
\end{abstract}
\tableofcontents %
\listoffigures %
%
%\listoftables %
%
\begin{acknowledgements}
Acknowledgements
\end{acknowledgements}
\pagestyle{thesis}
\newpage
\pagenumbering{arabic}

\input{sections/cosmo_intro.tex}
\input{sections/experiment.tex}
\input{sections/tension_intro.tex}
\input{sections/gg_split.tex}
\input{sections/emulators.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{unsrtnat}{}
\renewcommand{\baselinestretch}{1}
\normalsize

\clearpage
\newpage
\phantomsection%
\addcontentsline{toc}{chapter}{\numberline{}{Bibliography}}%
\bibliography{bib.bib}

\clearpage
\newpage

\appendix
\include{sections/appendixSdkr}
%\include{sections/appendixNOON}

\end{document}


