\chapter{Connecting to Experiments: Weak Lensing and the Cosmic Microwave Background}

\section{Weak Lensing}

In general, contemporary weak lensing surveys measure the light emmited from galaxies in two components: galaxy position and galaxy shear. In the following sections the theoretical base for cosmological weak lensing surveys are described. However, before discussing the two relevant observables, one needs to formalize the concept of weak lensing.

Weak lensing occurs when light travels with curved trajectories. Contrary to the purely Newtonian approach to gravity, in General Relativity light travels along geodesics with vanishing infinitessimal distances. Thus the time a photon has travelled is given by

\subsection{The Linear Matter Power Spectrum}

As we have seen before, we consider deviations from the FLRW metric using the scalar fields $\Phi$ and $\Psi$. In the absence of anisotropic stress, $\Phi=-\Psi$. Throughout time in the universe, we consider there to be three distinct epochs/regions:
\begin{itemize}
	\item The super-horizon region, or the region before inflation where radiation dominates
	\item The horizon-crossing region, or the region during inflation where the universe transfers from radiation to matter domination
	\item The sub-horizon region, or the region of after rapid inflation where matter dominates
\end{itemize}
Additionally, the formation of matter inhomogeneity comes from gravitational instability, the idea that primordial curvature perturbations in the super-horizon grew to form small matter density contrasts which grew in size after crossing the horizon. After the rate of inflation slowed, gravitational forces caused clusters of matter to become the galaxies and clusters we see today.

This gives us a blueprint for writing the matter power spectrum. We have the primordial curvature perturbation $\mathcal{R}$ which source the inital matter inhomogeneity that formed late-time large scale structure. Thus, we can write it as
\begin{equation}
	\mathcal{R} = \frac{5}{3}\Phi_{\text{large-scale}}(k,a_{\text{late}})
\end{equation} 
Additionally, we have a transfer function which describes the transition from radiation to matter domination, defined as
\begin{equation}
	T(k) = \frac{\Phi(k,a_{\text{late}})}{\Phi_{\text{large-scale}(k,a_{\text{late}})}}
\end{equation}
And we have a growth factor, defined as the linear growth of matter density contrast during matter domination, defined as
\begin{equation}
	D_+(a) = a \frac{\Phi(k,a)}{\Phi(k,a_{\text{late}})}
\end{equation}
Thus we can write the gravitational potential at all times as
\begin{equation}
	\Phi(k,a) = \frac{3}{5a}\mathcal{R}(k)T(k,a)D_+(a)
\end{equation}
This together with the Poisson equation for the matter density contrast field gives the equation
\begin{equation}
	\delta_m(k) = \frac{2k^2a}{3\Omega_mH_0^2}\Phi(k,a)
\end{equation}
\begin{equation}
	\delta_m(k) = \frac{2}{5}\frac{k^2}{\Omega_mH_0^2}\mathcal{R}(k)T(k,a)D_+(a)
\end{equation}
The power spectrum of an observable is defined as the two-point correlation function of the observable.
\begin{equation}
	P_\mathcal{O} = \langle \mathcal{O}\mathcal{O} \rangle
\end{equation}
Thus the linear matter power spectrum (because we restricted the matter density contrast field to linear order terms in the perturbation) is given by
\begin{equation}
	\begin{split}
		P_m^{L}(k,a) = \frac{8\pi^2}{25 \Omega_m^2 H_0^4} A_s D_+^2(a)T^2(k)\frac{k^{n_s}}{k_p^{(n_s-1)}}
	\end{split}
\end{equation}

\subsection{Galaxy Density}
In practice, we cannot measure the matter power spectrum. Dark matter prohibits direct measurement, thus one must determine observables which trace the matter distribution in the universe. The most natural option is to observe the galaxy distribution. However, large scale structures and galaxies provide many sources for non-linearities in the power spectrum and result in a non-trivial relation between the two. By taking certain data selection criteria, one can cut out the non-linear contributions. The relation between the galaxy and matter power spectra are then related by the linear bias factor.
\begin{equation}
	\begin{split}
		\delta_g(k) &= b_1 \delta_m(k) \\
		P_g(k,a) &= (b_1)^2P_m^L(k,a)
	\end{split}
\end{equation}
Realistically, in galaxy surveys we do not have access to exact reshifts for each observed galaxy. Rather, an image is taken of the sky in multiple color bands from which the redshift is determined. This procedure is called photometric redshift. To account for this, one can attempt to redefine the galaxy power spectrum. First, we define the distribution of distances as
\begin{equation}
	W(\chi) = \frac{1}{N_g}\frac{dN_g}{d\chi}
\end{equation}
The procedure for determining this distribution will be discussed in section (section for dz and IA). This allows us to simply write the projected density contrast as
\begin{equation}
	\Delta_g = \int\limits_0^\infty W(\chi)\delta_g(x,\tau)d\chi
\end{equation}
To complete this discussion, we will make some approximations. We will only consider small scales for which $\sin(\theta)\sim \theta \sim 1/l$ where $l$ is the Fourier conjugate of $\theta$. Second is the limber approximation, which states that, as long as the galaxy power spectrum $P_g$ slowly varies over the region $\Delta k \sim 1/l\chi$, we approximate is as a constant. These approximations greatly simplify the integrals required when computing power spectrum of Fourier transformed $\Delta_g$. This power spectrum is called the angular power spectrum. Taking into account the approximations above and computing the fourier transfrom, one finds an angular power spectrum of
\begin{equation}
 	C_g(l) =\int \frac{1}{\chi^2}W^2(\chi)P_g(k=(l+1/2)/\chi,\eta) d\chi
\end{equation} 
% One of the observables considered in weak lensing surveys is galaxy position. In this case, various physical effects will alter the observed redshift for a given galaxy, thus affecting our distance measurements. We will consider the following two:
% \begin{itemize}
% 	\item Inaccurate modelling of the distance-redshift relation. This involves incorrect calibration of $a(t)$, the scale factor. This manifests as a correction to the comoving distance, denoted $\delta\chi$.
% 	\item Additional redshift doppler shifts. This manifests as a correction to the reshift-distnace relation involving the galaxy's perculiar velocity along the line of sight given by
% 	\begin{equation}
% 	 	1+z = \frac{1}{a}(1+u_\parallel) \,
% 	\end{equation} 
% 	To linear order, the correction is given by $\partial_{u_\parallel} x|_{u_\parallel=0} u_{\parallel} = \frac{u_\parallel}{aH} $
% \end{itemize}
% Thus the total correction is
% \begin{equation}
% 	x_{\text{obs}} = x + \left( \frac{u_\parallel}{aH} + \delta\chi \right)\hat{n}
% \end{equation}
% Naturally, we can convert between the observed and true positions via the jacobian of the transformation.
% \begin{equation}
% 	\begin{split}
% 		d^3x_{\text{obs}} &= x_{\text{obs}}^2dx_{\text{obs}}d\Omega \\
% 		d^3x &= x^2dxd\Omega \\
% 		J &= \frac{dx}{dx_{\text{obs}}} \frac{x^2}{x^2_{\text{obs}}} \\
% 		\frac{dx}{dx_{\text{obs}}} &= 1 - \frac{d\delta\chi}{dx}\frac{dx}{dx_\text{obs}} - \frac{1}{aH}\frac{\partial u_\parallel}{dx}\frac{dx}{dx_{\text{obs}}} \\
% 		\frac{dx}{dx_{\text{obs}}} &= \left( 1 + \frac{d\delta\chi}{dx} + \frac{1}{aH}\frac{\partial u_\parallel}{dx} \right)^{-1} \\
% 		J &= \left( 1 + \frac{d\delta\chi}{dx} + \frac{1}{aH}\frac{\partial u_\parallel}{dx} \right)^{-1} 
% 			\left( x + \frac{\delta\chi}{x} + \frac{1}{aHx}\frac{\partial u_\parallel}{\partial x} \right)^{-2}
% 	\end{split}
% \end{equation}
% \begin{equation}
% 	\frac{d}{dx}\delta\chi = \frac{d\delta\chi}{dz}\frac{dz}{dx} = H\left(\frac{1}{H}-\frac{1}{H_{\text{fid}}}\right) = -\frac{1}{H_{\text{fid}}} \delta H
% \end{equation}
% The observed galaxy number density is related to the mean number density by the galaxy density contrast field
% \begin{equation}
% 	n = \bar n (1+\delta(x))
% \end{equation}




\subsection{Galaxy Shear}
% As a reminder, Christoffel symbols in perturbed spacetime to first order in the perturbations $\Psi$ and $\Phi$
% \begin{equation}
% 	\begin{split}
% 		\Gamma^{0}_{00} =& \dot\Psi \\
% 		\Gamma^{i}_{00} =& \frac{1}{a^2}\partial^i\Psi \\
% 		\Gamma^{i}_{j0} =& \delta_{ij}(H+\dot\Phi) \\
% 		\Gamma^{0}_{ij} =& \delta_{ij}a^2(H+2H(\Phi-\Psi)+\dot\Phi) \\
% 		\Gamma^{i}_{jk} =& [\delta_{ij}\partial_k + \delta_{ik}\partial_j - \delta_{jk}\partial_i]\Phi
% 	\end{split}
% \end{equation}
Photons move on null geodesics, thus $d\chi = -d\eta$, where $\eta$ is the conformal time and $\chi$ is the conformal distance. This tells us that $dx^i/d\chi = -dx^i/d\eta = -a dx^i/dt = -a dx^i/d\lambda d\lambda/dt = -a/P^0 P^\mu = -a / E * p/a \hat{p}^i = -\hat{p}^i$ (for photons and other massless particles).
Thus, under lensing, the difference between the observed and true positions within the source plane is given by
\begin{equation}
	\chi\mathbf{\theta}^i = x_\perp^i = -\int\limits_0^\chi \hat{p}^i_\perp(\chi'') d\chi''
\end{equation}
Our goal is to write this as a function of the potential only. Thus we look to compute $d\hat{p}^i/dt$.
\begin{equation}
	\begin{split}
		\frac{d\hat{p}^i}{dt} %=& \frac{d}{dt}\frac{1}{p}p^i \\
		=& \frac{1}{p^2}(\dot{p}^i p - \dot{p}p^i)
	\end{split}
\end{equation}
By computing $\dot p^i$, we can find $\dot p$ as well.
\begin{equation}
	\begin{split}
		\frac{dp^i}{d\lambda} %=& \frac{d}{d\lambda}\left(aP^i(1+\Phi)\right) \\
		=& P^i\frac{d}{d\lambda}(a(1+\Phi)+(1+\Phi)a\frac{d}{d\lambda}P^i
	\end{split}
\end{equation}
We have $d/d\lambda = P^\mu \partial_\mu$, so the first term becomes
\begin{equation}
	\begin{split}
		P^i\frac{d}{d\lambda}(a(1+\Phi)) %&= P^iP^\mu \partial_\mu(a(1+\Phi)) \\
		%&= P^iP^0(\dot a + a\dot\Phi + \dot a \Phi) + a P^iP^j \partial_j\Phi \\
		&= a P^i (P^0(H+\dot\Phi +H\Phi) + P^j\partial_j\Phi) \\
	\end{split}
\end{equation}
The second term is evaluated by using the geodesic equation up to first order in the perturbations.
\begin{equation}
	\begin{split}
		\frac{d}{d\lambda}P^i &= -\Gamma^i_{\mu\nu}P^\mu P^\nu \\
		%&= -(P^0)^2\frac{1}{a^2}\partial^i\Psi - 2\delta^{i}_{j}(H+\dot\Phi)P^0P^j + 2P^iP^j\partial_j\Phi - P^jP_j\partial^i\Phi \\
		%&= -E(1-\Psi)\left( \frac{E}{a^2} \partial^i\Psi + \frac{1}{a}2p^i(1-\Phi)(H+\dot\Phi) - \frac{2}{a^2E}p^ip^j\partial_j\Phi + \frac{p^2}{a^2E} \partial^i\Phi \right) \\
		&= -E\left( \frac{E}{a^2} \partial^i\Psi + \frac{1}{a}2p^i(1-\Phi-\Psi)(H+\dot\Phi) - \frac{2}{a^2E}p^ip^j\partial_j\Phi + \frac{p^2}{a^2E} \partial^i\Phi \right)
	\end{split}
\end{equation}
Plugging everything back in, and only keeping terms to linear order, we get
\begin{equation}
	\begin{split}
		%=& aP^i(H+\dot\Phi+H\Phi) + aP^iP^j\partial_j\Phi \\
		%& - \frac{E}{a}\partial^i\Psi - 2p^i(H+\dot\Phi) + \frac{2}{aE}p^ip^j \partial_j\Phi - \frac{p^2}{aE}\partial^i\Phi \\
		\frac{dp^i}{dt} %=& p^i(1-\Phi)(H+\dot\Phi+H\Phi) + (1-\Phi)^2p^ip^j\partial_j\Phi \\
		%& - \frac{E}{a^2}\partial^i\Psi -\frac{2p^i}{a}(H+\dot\Phi) + \frac{2}{a^2E}p^ip^j \partial_j\Phi - \frac{p^2}{a^2E}\partial^i\Phi \\
		%=& p^i(H+\dot\Phi) + \frac{p^ip^j}{aE}\partial_j\Phi \\
		%& - \frac{E}{a}\partial^i\Psi -2p^i(H+\dot\Phi) + \frac{2}{aE}p^ip^j \partial_j\Phi - \frac{p^2}{aE}\partial^i\Phi \\
		=& -p^i(H+\dot\Phi)- \frac{1}{aE}\left(p^ip^j\partial_j\Phi - p^2\partial^i\Phi\right) -\frac{E}{a}\partial^i\Psi
	\end{split}
\end{equation}
The time derivative of the modulus of the momentum, $p$, is then given by
\begin{equation}
	\begin{split}
		\frac{dp}{dt} =& \frac{d}{dt}\sqrt{\delta_{ij}p^ip^j} = \frac{1}{p}\delta_{ij}\dot p^i p^j \\
		%=& -p(H+\dot\Phi) - \frac{p}{aE}(p^i\partial_i\Phi - pp^i\partial^i\Phi) - p^i\frac{E}{ap}\partial^i\Psi \\
		=& -p(H+\dot\Phi) - \hat{p}_i\frac{E}{a}\partial^i\Psi
	\end{split}
\end{equation}
Now, putting everything together, plugging these results into eq () gives
\begin{equation}
	\begin{split}
		\frac{d\hat{p}^i}{dt} %=& -\hat{p}^i(H+\dot\Phi)- \frac{1}{aEp}\left(p^ip^j\partial_j\Phi - p\partial^i\Phi\right) -\frac{E}{ap}\partial^i\Psi + \hat{p}^i\left( (H+\dot\Phi) + \hat{p}_i\frac{E}{ap^2}\partial^i\Psi \right) \\
		=&  \frac{E}{ap}\left( \frac{p^2}{E^2}\partial_j\Phi - \partial_j\Psi \right)(\delta^{ij}-\hat{p}^i\hat{p}^j)
	\end{split}
\end{equation} 
Going back to the photon scenario, where $E=p$, this simplifies to
\begin{equation}
	\frac{d\hat{p}^i}{dt} = \frac{1}{a}\left( \partial_j\Phi - \partial_j\Psi \right)(\delta^{ij}-\hat{p}^i\hat{p}^j)
\end{equation}
Interestingly, $\delta^{ij}-\hat p^i \hat p^j$, is the projection operator of the $j$-th direction onto the $i$-th. So, if we orient our axes so that the photon approximately travels along the $j$-th direction, we can sum of $i$ and $k$ so that the orthogonal components of the momentum are evaluated by
\begin{equation}
	\frac{d\hat{p}^i_\perp}{d\chi} = - a \frac{d\hat{p}^i_\perp}{dt} = -\partial_i(\Phi-\Psi)
\end{equation}
In the absence of anisotropic stress (like in the late universe where weak lensing is measured), $\Phi=-\Psi$ and 
\begin{equation}
	\frac{d\hat{p}^i_\perp}{d\chi} -2\partial_i\Phi
\end{equation}
Integrating this equation gives us
\begin{equation}
	\hat p_\perp (\chi'') = -2\int\limits_0^{\chi''}\partial_i\Phi(\chi') d\chi' + C_i
\end{equation}
where the constant $C_i$ comes from integrating a derivative, where it is only defined up to a constant. Now, plugging back in we have
\begin{equation}
	\mathbf{\theta}^i = \frac{2}{\chi}\int\limits_0^\chi \int\limits_0^{\chi''}\partial_i\Phi(\chi') d\chi' d\chi'' - C_i
\end{equation}
In the absence of lensing, this should reduce to $\mathbf{\theta}^i = \theta_0$, the true source position. Hence $C^i=\theta^i_0$ and the total deflection angle is given by
\begin{equation}
	\begin{split}
		\Delta\theta^i =& \frac{2}{\chi}\int\limits_0^\chi \int\limits_0^{\chi''}\partial_i\Phi(\chi') d\chi' d\chi''\\
		=& \frac{2}{\chi}\int\limits_0^\chi \partial_i\Phi(\chi') (\chi-\chi') d\chi'
	\end{split}
\end{equation}
Finally, since $x^i=\chi\theta^i$, we have $\partial_i = \partial_{\theta^i}/\chi$. Now we introduce the \textit{lensing potential} defined by
\begin{equation}
	\begin{split}
		\Delta\theta^i =& \frac{\partial}{\partial \theta^i}\psi_L(\theta) \\
		\psi_L(\theta) =& 2\int\limits_0^\chi \frac{1}{\chi'}\Phi(\chi') \left(1-\frac{\chi'}{\chi}\right) d\chi'
	\end{split}
\end{equation}
This result is great, but it's not very experimentally illuminating. Similar to the galaxy vs matter power spectrum, the absence of the observation of dark matter means we cannot reliably determine the potential $\Phi$. Again, an attempt to relate the lensing potential to some property of galaxies can give us a measureable quantity. Lensing distorts the shape of the galaxies (assuming they are all perfectly elliptical). Define the galaxy shape tensor by
\begin{equation}
	q_{ij} \equiv \frac{1}{F}\int I\theta^i\theta^j d\theta^i d\theta^j
\end{equation}
This integral is symmetric under $i \leftrightarrow j$. Writing this as a matrix gives
\begin{equation}
	q_{ij} = \frac{1}{2} \text{tr}(q) \left( 
	\begin{array}{cc}
	1+\epsilon_1 & \epsilon_2 \\
	\epsilon_2 & 1-\epsilon_1
	\end{array}
	\right)
\end{equation}
Shape distortion specifically occurs when the deflection angle is not constant across a galaxy. Thus we can look at the tranformation tensor as
\begin{equation}
	A_{ij} \equiv \frac{\partial \theta^i_S}{\partial\theta^j} = \left( 
	\begin{array}{cc}
	1+\kappa - \gamma_+ & -\gamma_\times \\
	-\gamma_\times & 1+\kappa+\gamma_+
	\end{array}
	\right) = \delta_{ij}+\psi_{ij}
\end{equation}
Within this tensor, the $\gamma_+$ and $\gamma_-$ represent shear and $\kappa$ represents magnification and changes in galaxy light flux ($F$). The $E$-mode is given by
\begin{equation}
	E(l) = \left( \frac{l^il^j}{l^2} - \frac{1}{2}\delta^{ij} \right)(-\psi^{ij}(l))
\end{equation}
Which relates to the shear components by
\begin{equation}
	\left(
	\begin{array}{c}
	\gamma_+ \\
	\gamma_\times
	\end{array}
	\right)
	=
	\left(
	\begin{array}{c}
	\cos(2\alpha_l) \\
	\sin(2\alpha_l)
	\end{array}
	\right)
	E(l)
\end{equation}
This connects the observable shear to the theoretical lensing potential.
\subsection{Weak Lensing Correlation Functions}
In the previous two sections, we have derived two observables directly connected to cosmological theories. By observing the galaxies we can determine $\Omega_b$, $A_s$, $n_s$ and $H_0$. By observing shear we can determine $\Omega_m$. The details for how to determine these parameters will be discuss in chapter (chapter). Together these two can fully determine the values of the 5 $\Lambda$CDM parameters. We are still not ready, since these two quantities themselves are not observables. Using these two quantities, however, one can construct three different two-point correlation functions which are observable.

The first correlation function is the autocorrelation of galaxy density, called galaxy clustering. Given a density in a particular angular position, the correlation function can be interpreted as the likelihood of observing a similar density at another close position. The galaxy clustering correlation function is given by
\begin{equation}
	w_g(\theta) = \mathcal{F}(C_g(l)) = \frac{1}{2\pi}\int\limits_0^\infty l C_g(l) J_0(l\theta)dl
\end{equation}
Where $J_0$ is the 0-th Bessel function and $\mathcal{F}$ denotes the fourier transform.

Next is the shear autocorrelation. In this case, there are two components of shear: the tangential shear $\gamma_+$ and the cross shear $\gamma_\times$. The cross correlation $\langle \gamma_+ \gamma_\times \rangle$ vanishes, so we are left with combinations
\begin{equation}
	\begin{split}
		\langle \gamma_+\gamma_+ \rangle \pm \langle \gamma_\times \gamma_\times \rangle = \xi_\pm
	\end{split}
\end{equation}
\begin{equation}
	\xi_{+,-} = \frac{1}{2\pi}\int\limits_0^\infty l C_{EE}(l) J_{0,4}(l\theta)C_{EE}(l)dl
\end{equation}

Finally, we have the cross correlation between galaxy density and tangential shear. First, the angular power spectrum $C_{gE}(l)$ is given by
\begin{equation}
	C_{gE}(l) = \frac{3}{2}\Omega_mH_0^2 \int\limits_0^\infty \frac{1}{\chi a(\chi)}W_g(\chi)P_g(k,\eta) d\chi
\end{equation}
\begin{equation}
	\gamma = \langle \Delta_g\gamma_+\rangle = -\frac{1}{2\pi}\int\limits_0^\infty l J_2(l\theta)C_{gE}(l)dl
\end{equation}
The correlation function $\langle \Delta_g\gamma_\times\rangle$ vanishes in an isotropic universe.

\subsection{Systematics}
Up to now, we have neglected any systematics of the weak lensing observables. There are two distinct types of systematics that will be discussed in this section: The ones which affect the calculation of the correlation functions, and ones which affect the correlation functions after they have been computed. The latter type are usually referred two as `fast' parameters, and we will see why in the next chapter. In the meantime here is an explanation of the relevant systematics.
\subsubsection{Photometric Redshift Uncertainty}
As we discussed in the galaxy density section, to perform the computation one must determine the redshift distrubution $W(\chi)$. This is a highly non-trivial task. To model the uncertainty in this distribution, one applies a constant shift to the redshift of observed galaxies.
\begin{equation}
	n_g(z) \mapsto n_g(z+\Delta z)
\end{equation}
The shift $\Delta z$ must be determined before one can compute the galaxy clustering and galaxy lensing correlation functions.
\subsubsection{Intrinsic Alignment of Galaxies}
As can be seen in figure (fig), galaxy shapes are not completely random. In the absence of lensing, there is still some non-zero correlation between the shapes. This is referred to as Intrinsic Alignment. There are two models of intrinsic alignment that will be considered in this thesis: the Tidal Alignment Tidal Torquing (TATT) model and the Non-linear Alignment (NLA) model. The TATT depends on the tidal tensor $s_{ij}$ which contains the information about tidal forces due to non-uniform gravitational fields. Tidal forces are a source of intrinsic alignment and shear (figure). The model is given by
\begin{equation}
	\gamma_{ij}^I = \underbrace{C_1s_{ij}}_{\text{tidal alignment}}+
	\underbrace{b_{TA}C_1(\delta_m\times s_{ij})}_{\text{density weighting}}+
	\underbrace{C_2\left( s_i^ks_{kj}-\frac{1}{3}\delta_{ij}s^2 \right)}_{\text{tidal torquing}}
\end{equation}
Where the coefficients $C_a$ are given by
\begin{equation}
	\begin{split}
		C_1 =& -\frac{A_1\bar{C}\Omega_m}{aG(z)}\left(\frac{1+z}{1+z_0}\right)^{\eta_1} \\
		C_2 =& 5\frac{A_2\bar{C}\Omega_m}{(aG(z))^2}\left(\frac{1+z}{1+z_0}\right)^{\eta_2}
	\end{split}
\end{equation}
The TATT model reduces to the NLA model when $A_2=0=b_{TA}$
\subsubsection{Galaxy Bias}
As discussed in the galaxy clustering section, we relate the galaxy and matter density by a bias parameter $b$. However we can pick a ficuial value to relate the two
\begin{equation}
	\delta_{g,\text{fid}} = b_{\text{fid}}\delta_m
\end{equation}
then consider a multiplicative deviation
\begin{equation}
	\delta_g = (b\cdot b_{\text{fid}})\delta_m = b\delta_{g,\text{fid}}
\end{equation}
Thus, the parameter $b$ can be found to account for different galaxy biases. 

Galaxy bias is the first of many `fast' parameters. Since the bias is constant, it can be pulled out of the integral for the correlation functions, meaning it can be applied after the correlation function has been computed. The affected correlation functions are the galaxy clustering and galaxy-galaxy lensing.
\begin{equation}
	\begin{split}
		w^i(\theta) =& b_{(i)}^2w^i(\theta) \\
		\gamma^i_t(\theta) =& b_{(i)}\gamma_t^i(\theta)
	\end{split}
\end{equation}
where $i$ here denotes the redshift bin
\subsubsection{Shear Calibration}
Observed shear is not a completely astrophysical phenomenon. For example, a major source of additional shear comes from the observation lense iteself. The exact methods for how this is handled (such as the point spread function) are beyond the scope of this thesis. This is usually modelled as a multiplicative correction to the shear componenets.
\begin{equation}
	\begin{split}
		\xi^{ij}_{\pm} =& (1+m^i)(1+m^j) \xi^{ij}_{\pm} \\
		\gamma^{ij}_t =& (1+m^j) \gamma_t^{ij}
	\end{split}
\end{equation}
where $i$ and $j$ denote the source and lens bin respectively. The multiplicative shear calibration is also a fast parameter.
\subsubsection{Point Mass}
\subsubsection{Lensing Magnification}
As discussed in (section), the intrinsic shape of galaxies also gets magnified. This effect is introduced as a parmeter $C_l^i$ and is fixed experimentally.

%Additionally, DES added a non-physical parameter called $X_{\text{lens}}$ to resolve an inconsistency between two lensing samples, RedMaGiC and MagLim. In this thesis, we only consider $X_{\text{\lens}}=1$

\section{CMB}
\subsection{TT,TE,EE Power Spcetra}
\subsection{Systematics}








